$Z^\prime$ mass limits and the naturalness of supersymmetry

### 2015-06-19 paper/e6ssm-tuning-paper.tex
Corrected several typos

$U(1)$ extensions of the MSSM provide a very elegant solution to this
$\mu$-problem \cite{Fayet:1977yc, Kim:1983dt, Suematsu:1994qm,
  Cvetic:1995rj, Cvetic:1996mf, Jain:1995cb, Nir:1995bu,
  Cvetic:1997ky} and also raise the Higgs mass with new $F$- and
$D$-terms.  Nonetheless, as was recently demonstrated in the context
of the exceptional supersymmetric standard model (E$_6$SSM)
\cite{King:2005jy, King:2005my, Athron:2010zz}, such models can still
suffer from naturalness problems with the mass of the new $Z^\prime$
associated with the break down of the new $U(1)$ appearing in the
electroweak symmetry breaking conditions at tree level
\cite{Athron:2013ipa}.  Despite this the constrained version of the
E$_6$SSM (cE$_6$SSM) \cite{Athron:2009ue, Athron:2009bs} was still
found to be significantly less tuned than the cMSSM.  Tree level
fine tuning from the $Z^\prime$ mass was also considered
previously \cite{Drees:1985js}.

@@

At the same time in $U(1)$ extensions, lowering the UV boundary scale
for the RGE evolution also allows even larger $F$-term contributions
to the Higgs mass, so long as one only requires $\lambda$, the
coupling between the Singlet Higgs, $S$ and the up- and down-type
Higgs bosons, $H_u$ and $H_d$, to remain perturbative up to the UV
scale and not the all the way up to the GUT scale.

@@

Demanding that the Higgs fields $H_1$, $H_2$ and the singlet $S$ acquire real
VEVs of the form
\begin{equation} \label{eq:E6vevs}
\langle H_d \rangle = \frac{1}{\sqrt{2}} \begin{pmatrix} v_1 \\ 0
\end{pmatrix} , \;
\langle H_u \rangle = \frac{1}{\sqrt{2}} \begin{pmatrix} 0 \\ v_2
\end{pmatrix} , \;
\langle S \rangle =\frac{s}{\sqrt{2}} ,
\end{equation}
at the physical minimum leads to the minimization conditions
\begin{subequations} \label{eq:E6EWSBConditions}
\begin{align}
&f_1 = m_{H_d}^2 v_1 + \frac{\lambda^2}{2} (v_2^2 + s^2) v_1-
\frac{\lambda A_\lambda}{\sqrt{2}} s v_2 - \frac{\bar{g}^2}{8}
(v_2^2 - v_1^2) v_1 + D_{H_d} v_1 + \frac{\partial \Delta V}{\partial v_1} = 0 ,
\label{eq:E6EWSBcondition1} \\
&f_2 = m_{H_u}^2 v_2 + \frac{\lambda^2}{2} (v_1^2 + s^2) v_2 -
\frac{\lambda A_\lambda}{\sqrt{2}} s v_1 + \frac{\bar{g}^2}{8}
(v_2^2 - v_1^2) v_2 + D_{H_u} v_2 + \frac{\partial \Delta V}{\partial v_2} = 0 ,
\label{eq:E6EWSBcondition2} \\
&f_3 = m_S^2 s + \frac{\lambda^2}{2} (v_2^2 + v_1^2) s -
\frac{\lambda A_\lambda}{\sqrt{2}} v_2 v_1 + D_S s +
\frac{\partial \Delta V}{\partial s} = 0 . \label{eq:E6EWSBcondition3}
\end{align}
\end{subequations}
The quantities $D_{H_d}$, $D_{H_u}$ and $D_S$ appearing above are $U(1)'$
$D$-term contributions that are absent in the MSSM and NMSSM and are given by
\begin{equation} \label{eq:E6Dterms}
D_\phi \equiv \frac{g_1'^2}{2} \left (Q_1 v_1^2 + Q_2 v_2^2 + Q_S s^2
\right ) Q_\phi .
\end{equation}
We also include these $U(1)'$ $D$-term contributions in the diagonalized stop
masses,
\begin{align}
m_{\tilde{t}_{1,2}}^2 &= \frac{1}{2} \bigg \{ m_{Q_3}^2 + m_{u_3}^2 +
\frac{\bar{g}^2}{8} \left ( v_1^2 - v_2^2 \right ) + D_Q + D_u + 2 m_t^2
\nonumber \\
& \quad {} \mp \sqrt{\left [ m_{Q_3}^2 + m_{u_3}^2 + \frac{1}{8} \left (g_2^2
- g_1^2 \right ) \left (v_1^2 - v_2^2 \right ) + D_Q - D_u \right ]^2 +
4 m_t^2 X_t^2} \bigg \} ,
\end{align}
where $m_t^2 = y_t^2 v_2^2 / 2$, $X_t = A_t -
\frac{\lambda s v_1}{\sqrt{2} v_2}$, $m_{Q_3}^2$, $m_{u_3}^2$ are soft
breaking scalar masses and $A_t$ is a soft trilinear coupling.  By definition
we take $m_{\tilde{t}_1}$ to correspond to the lighter of the two states.

@@

However, it is well known in the MSSM that radiative corrections can
significantly change (indeed, reduce) the fine tuning \cite{Cassel:2010px}.
It is therefore important when studying the fine tuning to include loop
corrections to the effective potential in the fine tuning measure.  To do so
it is most convenient to work in terms of the EWSB conditions
Eq.~(\ref{eq:E6EWSBConditions}), rather than Eq.~(\ref{eq:E6MZequation}).
The general procedure that we use is as follows (this method has also
previously been applied in the NMSSM, see for example
Ref.~\cite{Ellwanger:2011mu}).  For a model in which $m$ fields develop real
VEVs (e.g. $m=2$ in the MSSM, $m=3$ in the NMSSM and in the $E_6$
models considered), we require that the $m$ minimization conditions
\begin{equation} \label{eq:EWSBconditions}
f_1 = f_2 = \dotsb = f_m = 0
\end{equation}
continue to hold under an arbitrary variation in a model parameter
$p \rightarrow p + \delta p$, so that the variations $\delta f_i$ satisfy
\begin{equation} \label{eq:EWSBvariations}
\delta f_1 = \delta f_2 = \dotsb = \delta f_m = 0 .
\end{equation}
Each $f_i$ is a function of the VEVs $v_j$ and $l$ running parameters
$q_k$ evaluated at the scale of EWSB, $f_i = f_i(v_j , q_k)$.  Thus for
each $f_i$ we have that
\begin{equation} \label{eq:EWSBchainrule}
\delta f_i = \sum_{j=1}^m \frac{\partial f_i}{\partial v_j}
\frac{\partial v_j}{\partial p} + \sum_{k=1}^l \frac{\partial f_i}
{\partial q_k} \frac{\partial q_k}{\partial p} = 0 .
\end{equation}
The quantities $\frac{\partial f_i}{\partial v_j}$ are the elements of
the CP-even Higgs squared mass matrix $M_h^2$ of the model before
rotating into the mass eigenstate basis.  When evaluated for all $n$
model parameters, the above system of equations can be concisely
expressed as
\begin{equation} \label{eq:tuningsystem}
M_h^2 \begin{pmatrix}
\frac{\partial v_1}{\partial p_1} & \cdots &
\frac{\partial v_1}{\partial p_n} \\
\vdots & \ddots & \vdots \\
\frac{\partial v_m}{\partial p_1} & \cdots & \frac{\partial v_m}{\partial p_n}
\end{pmatrix} = -\begin{pmatrix}
\frac{\partial f_1}{\partial q_1} & \cdots & \frac{\partial f_1}
{\partial q_l} \\
\vdots & \ddots & \vdots \\
\frac{\partial f_m}{\partial q_1} & \cdots & \frac{\partial f_m}{\partial q_l}
\end{pmatrix}
\begin{pmatrix}
\frac{\partial q_1}{\partial p_1} & \cdots & \frac{\partial q_1}
{\partial p_n} \\
\vdots & \ddots & \vdots \\
\frac{\partial q_l}{\partial p_1} & \cdots & \frac{\partial q_l}{\partial p_n}
\end{pmatrix} .
\end{equation}
The quantities forming the first matrix on the right-hand side, along
with $M_h^2$, are easily calculated by differentiating the conditions
in Eq.~(\ref{eq:E6EWSBConditions}) with respect to the VEVs and the
running parameters.  The remaining derivatives $\partial q_k /\partial
p$ must be determined using the RGEs.  Once these have been obtained,
it is straightforward to solve for the $\partial v_i / \partial p$.
The sensitivities $\Delta_p$ are then simply linear combinations of
the $\partial v_i / \partial p$ and $\partial q_k / \partial p$.  The
effects of radiative corrections may be easily included by including
the Coleman-Weinberg potential contributions $\Delta V$ in the EWSB
conditions.  Here we use the one-loop corrections given in
Eq.~(\ref{Eq:DeltaV}).

@@

To illustrate this, in the left panel of Fig.~\ref{Fig:mssmstoptuning}
we show the variation in fine tuning for $M_X = 20$ TeV, $50$ TeV,
$100$ TeV and $10^{16}$ GeV when we scan over the stop masses and
mixing, with $500 \textrm{ GeV} \leq m_{Q_3} , m_{u_3} \leq 10
\textrm{ TeV}$ and $-3810\textrm{ GeV} \leq A_t \leq -20 \textrm{
  GeV}$.  The remaining parameters we fix such that at $M_{\rm SUSY}$
they have the values $\mu = -97.5$ GeV, $B = -84.8$ GeV, $M_1 = 92.1$
GeV, $M_2 = 95.9$ GeV, $M_3 = 352$ GeV, $A_b = -117.9$ GeV, $A_\tau =
-7.8$ GeV, $m_{L_i} = 400$ GeV, $m_{e_i} = 204$ GeV, $m_{Q_{1,2}} =
438$ GeV, $m_{u_{1,2}} = 436$ GeV and $m_{d_i} = 438$ GeV ($i = 1 , 2
, 3$).  Here we denote by $M_1$, $M_2$ and $M_3$ the soft gaugino
masses for $U(1)$, $SU(2)$ and $SU(3)$, while $A_b$ and $A_\tau$ are
soft trilinear couplings and the $m_\phi$ are soft scalar masses for
the indicated fields.  The soft bilinear coupling $B$ is defined such
that at tree level the mass of the CP-odd MSSM Higgs boson reads
$m_A^2 = 2 B \mu / \sin 2\beta$.  All off-diagonal couplings and
scalar masses are set to zero, as are the first and second generation
Yukawa couplings and soft trilinears.  Although we should stress that
making this choice will lead to a spectrum which is in conflict with
the LHC limits, doing so ensures that fine tuning due to the other
parameters is small, so that we avoid washing out the fine tuning
impact of the stops when the tuning is small\footnote{For models in
  which the spectrum is heavier, when the stop masses are small the
  fine tuning reaches a lower bound imposed by other heavier
  parameters.} as can be the case when the stop masses are less than 1
TeV.  Note that the Higgs mass is also allowed to vary in this scan,
as shown in the right panel of Fig.~\ref{Fig:mssmstoptuning}.  This
illustrates the tuning problem which people have been worrying about
since the discovery of the 125 GeV Higgs boson as we see that raising
the stop masses is also pushing up the Higgs mass, meaning that
heavier Higgs masses require more fine tuning.  However for a low
value of the UV scale this tuning is not so severe unless the stops
are very heavy, and a 125 GeV Higgs can be obtained without much
tuning in this unrealistic case where we have minimized other sources
of tuning.  On the other hand the tuning becomes more severe as we
increase the cutoff such that for $M_X = 10^{16}$ GeV a lightest stop
masses of $1-3$ TeV can result in a fine tuning of $\approx 100 -
1000$ and the minimum tuning we find\footnote{Note that in the
  calculation of the Higgs mass there is a significant theoretical
  error, even with leading two-loop corrections, which should be
  considered when thinking about what the results imply for the
  minimum fine tuning in the model consistent with the recent
  discovery of a 125 GeV Higgs.} for a $125$ GeV Higgs is $\approx
200$, as shown in Fig.~\ref{Fig:mssmstoptuning}.

@@

MSSM benchmark BM2 (dash-dotted in top panel, individual sensitivities in
bottom left panel) is designed to be similar to E$_6$SSM BM2, for a reasonable
comparison.  However from the individual sensitivities one can see that the
behaviour is quite similar to MSSM BM1, though in this case the
$\Delta_{m_{H_u}^2}$ becomes the largest tuning at a higher $M_X$ and does not
reach such large values, since more of the tuning is from the mixing in this
case.

@@

Prior to stringent experimental constraints on the mass of the
lightest Higgs boson and squarks in supersymmetric models, a simple
picture of a natural SUSY model emerged from theoretical reasoning,
with soft masses set to similar values at the GUT scale through local
gravity interactions with the hidden sector.  Through the use of
renormalization group running one can then see that at the EW scale
the stops enter the EWSB condition for $M_Z$, therefore it was
expected that these masses should not be much bigger than $100$ GeV.
However to disturb this elegant picture first LEP placed constraints
on the Higgs mass, requiring it to be above 114.4 GeV
\cite{Barate:2003sz, Schael:2006cr}, which already introduced
significant tuning for constrained models, since heavy stops are
required to raised the lightest Higgs mass above it's tree level
bound. Then recently this problem got much worse since the LHC measured the
Higgs mass to be around $125$ GeV.

@@

In contrast we find that in $U(1)$ extensions of the MSSM there is an
additional way to constrain the naturalness of the models, which is
through the $Z^\prime$ mass limit.  We find when we impose a low
cutoff of $20$ TeV for setting the soft masses, the lowest tuning in
the E$_6$SSM is compatible with a $Z^\prime$ mass of $2.5$ TeV was
$\Delta \approx 121$, while if the LHC run II can place a limit of
$4.5$ TeV on $M_Z^\prime$ then the tuning would be approximately $394$.
By comparison the current situation in the MSSM only requires a tuning
of around $38$.  This should be interpreted as saying that in
the most conservative limits one can place on naturalness in these
models, the tuning in the E$_6$SSM is worse.  However if there are no
charginos below $700$ GeV then the situation in the two models would
be the same.

@@

When this trace is non-zero it will also induce gauge kinetic mixing
to be generated during RGE evolution and this is the case in the
models we consider here\footnote{In $E_6$ inspired models with only
complete $27$-plet matter multiplets this trace would vanish.
However since we assume some incomplete multiplets so that our
models are consistent with gauge coupling unification this trace
doesn't vanish.}.  However when these models are evolved down from
the GUT scale the radiatively generated gauge kinetic mixing gives an
off diagonal gauge coupling, $g_{11}$, of just $\approx 0.02$ at the
electroweak scale \cite{King:2005jy} and so it does not play a large
role.  At the same time if gauge kinetic mixing is included the RGE
expressions become very large and unmanageable, so we neglected the
gauge kinetic mixing by setting the SARAH flag \texttt{NoU1Mixing} to true.

@@

& {} - 3 \left ( 3 y_t^4 + 3 y_b^4 + 2 y_t^2 y_b^2 + y_\tau^2 \right ) +

@@

\right ) + 4 y_t^2 y_b^2 \left ( m_{H_u}^2 + m_{H_d}^2 + m_{Q_3}^2 +

>>>>

$U(1)$ extensions of the MSSM provide a very elegant solution to this
$\mu$-problem \cite{Fayet:1977yc, Kim:1983dt, Suematsu:1994qm,
  Cvetic:1995rj, Cvetic:1996mf, Jain:1995cb, Nir:1995bu,
  Cvetic:1997ky} and also raise the Higgs mass with new $F$- and
$D$-terms.  Nonetheless, as was recently demonstrated in the context
of the exceptional supersymmetric standard model (E$_6$SSM)
\cite{King:2005jy, King:2005my, Athron:2010zz}, such models can still
suffer from naturalness problems with the mass of the new $Z^\prime$
associated with the break down of the new $U(1)$ appearing in the
electroweak symmetry breaking (EWSB) conditions at tree level
\cite{Athron:2013ipa}.  Despite this the constrained version of the
E$_6$SSM (cE$_6$SSM) \cite{Athron:2009ue, Athron:2009bs} was still
found to be significantly less tuned than the cMSSM.  Tree level
fine tuning from the $Z^\prime$ mass was also considered
previously \cite{Drees:1985js}.

@@

At the same time in $U(1)$ extensions, lowering the UV boundary scale
for the RGE evolution also allows even larger $F$-term contributions
to the Higgs mass, so long as one only requires $\lambda$, the
coupling between the Singlet Higgs, $S$ and the up- and down-type
Higgs bosons, $H_u$ and $H_d$, to remain perturbative up to the UV
scale and not all of the way up to the GUT scale.

@@

Demanding that the Higgs fields $H_1$, $H_2$ and the singlet $S$ acquire real
VEVs of the form
\begin{equation} \label{eq:E6vevs}
\langle H_d \rangle = \frac{1}{\sqrt{2}} \begin{pmatrix} v_1 \\ 0
\end{pmatrix} , \;
\langle H_u \rangle = \frac{1}{\sqrt{2}} \begin{pmatrix} 0 \\ v_2
\end{pmatrix} , \;
\langle S \rangle =\frac{s}{\sqrt{2}} ,
\end{equation}
at the physical minimum leads to the minimization conditions
\begin{subequations} \label{eq:E6EWSBConditions}
\begin{align}
&f_1 = m_{H_d}^2 v_1 + \frac{\lambda^2}{2} (v_2^2 + s^2) v_1-
\frac{\lambda A_\lambda}{\sqrt{2}} s v_2 - \frac{\bar{g}^2}{8}
(v_2^2 - v_1^2) v_1 + D_{H_d} v_1 + \frac{\partial \Delta V}{\partial v_1} = 0 ,
\label{eq:E6EWSBcondition1} \\
&f_2 = m_{H_u}^2 v_2 + \frac{\lambda^2}{2} (v_1^2 + s^2) v_2 -
\frac{\lambda A_\lambda}{\sqrt{2}} s v_1 + \frac{\bar{g}^2}{8}
(v_2^2 - v_1^2) v_2 + D_{H_u} v_2 + \frac{\partial \Delta V}{\partial v_2} = 0 ,
\label{eq:E6EWSBcondition2} \\
&f_3 = m_S^2 s + \frac{\lambda^2}{2} (v_2^2 + v_1^2) s -
\frac{\lambda A_\lambda}{\sqrt{2}} v_2 v_1 + D_S s +
\frac{\partial \Delta V}{\partial s} = 0 . \label{eq:E6EWSBcondition3}
\end{align}
\end{subequations}
The quantities $D_{H_d}$, $D_{H_u}$ and $D_S$ appearing above are $U(1)'$
$D$-term contributions that are absent in the MSSM and NMSSM and are given by
\begin{equation} \label{eq:E6Dterms}
D_\phi \equiv \frac{g_1'^2}{2} \left (Q_1 v_1^2 + Q_2 v_2^2 + Q_S s^2
\right ) Q_\phi .
\end{equation}
We also include these $U(1)'$ $D$-term contributions in the diagonalized stop
masses,
\begin{align}
m_{\tilde{t}_{1,2}}^2 &= \frac{1}{2} \bigg \{ m_{Q_3}^2 + m_{u_3}^2 +
\frac{\bar{g}^2}{8} \left ( v_1^2 - v_2^2 \right ) + D_Q + D_u + 2 m_t^2
\nonumber \\
& \quad {} \mp \sqrt{\left [ m_{Q_3}^2 - m_{u_3}^2 + \frac{1}{8} \left (g_2^2
- g_1^2 \right ) \left (v_1^2 - v_2^2 \right ) + D_Q - D_u \right ]^2 +
4 m_t^2 X_t^2} \bigg \} ,
\end{align}
where $m_t^2 = y_t^2 v_2^2 / 2$, $X_t = A_t -
\frac{\lambda s v_1}{\sqrt{2} v_2}$, $m_{Q_3}^2$, $m_{u_3}^2$ are soft
breaking scalar masses and $A_t$ is a soft trilinear coupling.  By definition
we take $m_{\tilde{t}_1}$ to correspond to the lighter of the two states.

@@

However, it is well known in the MSSM that radiative corrections can
significantly change (indeed, reduce) the fine tuning \cite{Cassel:2010px}.
It is therefore important when studying the fine tuning to include loop
corrections to the effective potential in the fine tuning measure.  To do so
it is most convenient to work in terms of the EWSB conditions
Eq.~(\ref{eq:E6EWSBConditions}), rather than Eq.~(\ref{eq:E6MZequation}).
The general procedure that we use is as follows (this method has also
previously been applied in the NMSSM, see for example
Ref.~\cite{Ellwanger:2011mu}).  For a model in which $m$ fields develop real
VEVs (e.g. $m=2$ in the MSSM, $m=3$ in the NMSSM and in the $E_6$
models considered), we require that the $m$ minimization conditions
\begin{equation} \label{eq:EWSBconditions}
f_1 = f_2 = \dotsb = f_m = 0
\end{equation}
continue to hold under an arbitrary variation in a model parameter
$p \rightarrow p + \delta p$, so that the variations $\delta f_i$ satisfy
\begin{equation} \label{eq:EWSBvariations}
\delta f_1 = \delta f_2 = \dotsb = \delta f_m = 0 .
\end{equation}
Each $f_i$ is a function of the VEVs $v_j$ and $l$ running parameters
$q_k$ evaluated at the scale of EWSB, $f_i = f_i(v_j , q_k)$.  Thus for
each $f_i$ we find that
\begin{equation} \label{eq:EWSBchainrule}
\sum_{j=1}^m \frac{\partial f_i}{\partial v_j}
\frac{\partial v_j}{\partial p} + \sum_{k=1}^l \frac{\partial f_i}
{\partial q_k} \frac{\partial q_k}{\partial p} = 0 .
\end{equation}
The quantities $\frac{\partial f_i}{\partial v_j}$ are the elements of
the CP-even Higgs squared mass matrix $M_h^2$ of the model before
rotating into the mass eigenstate basis.  When evaluated for all $n$
model parameters, the above system of equations can be concisely
expressed as
\begin{equation} \label{eq:tuningsystem}
M_h^2 \begin{pmatrix}
\frac{\partial v_1}{\partial p_1} & \cdots &
\frac{\partial v_1}{\partial p_n} \\
\vdots & \ddots & \vdots \\
\frac{\partial v_m}{\partial p_1} & \cdots & \frac{\partial v_m}{\partial p_n}
\end{pmatrix} = -\begin{pmatrix}
\frac{\partial f_1}{\partial q_1} & \cdots & \frac{\partial f_1}
{\partial q_l} \\
\vdots & \ddots & \vdots \\
\frac{\partial f_m}{\partial q_1} & \cdots & \frac{\partial f_m}{\partial q_l}
\end{pmatrix}
\begin{pmatrix}
\frac{\partial q_1}{\partial p_1} & \cdots & \frac{\partial q_1}
{\partial p_n} \\
\vdots & \ddots & \vdots \\
\frac{\partial q_l}{\partial p_1} & \cdots & \frac{\partial q_l}{\partial p_n}
\end{pmatrix} .
\end{equation}
The quantities forming the first matrix on the right-hand side, along
with $M_h^2$, are easily calculated by differentiating the conditions
in Eq.~(\ref{eq:E6EWSBConditions}) with respect to the VEVs and the
running parameters.  The remaining derivatives $\partial q_k /\partial
p$ must be determined using the RGEs.  Once these have been obtained,
it is straightforward to solve for the $\partial v_i / \partial p$.
The sensitivities $\Delta_p$ are then simply linear combinations of
the $\partial v_i / \partial p$ and $\partial q_k / \partial p$.  The
effects of radiative corrections may be easily included by including
the Coleman-Weinberg potential contributions $\Delta V$ in the EWSB
conditions.  Here we use the one-loop corrections given in
Eq.~(\ref{Eq:DeltaV}).

@@

To illustrate this, in the left panel of Fig.~\ref{Fig:mssmstoptuning}
we show the variation in fine tuning for $M_X = 20$ TeV, $50$ TeV,
$100$ TeV and $10^{16}$ GeV when we scan over the stop masses and
mixing, with $500 \textrm{ GeV} \leq m_{Q_3} , m_{u_3} \leq 10
\textrm{ TeV}$ and $-3810\textrm{ GeV} \leq A_t \leq -20 \textrm{
  GeV}$.  The remaining parameters we fix such that at $M_{\rm SUSY}$
they have the values $\mu = -97.5$ GeV, $B = -84.8$ GeV, $M_1 = 92.1$
GeV, $M_2 = 95.9$ GeV, $M_3 = 352$ GeV, $A_b = -117.9$ GeV, $A_\tau =
-7.8$ GeV, $m_{L_i} = 400$ GeV, $m_{e_i} = 204$ GeV, $m_{Q_{1,2}} =
438$ GeV, $m_{u_{1,2}} = 436$ GeV and $m_{d_i} = 438$ GeV ($i = 1 , 2
, 3$).  Here we denote by $M_1$, $M_2$ and $M_3$ the soft gaugino
masses for $U(1)$, $SU(2)$ and $SU(3)$, while $A_b$ and $A_\tau$ are
soft trilinear couplings and the $m_\phi$ are soft scalar masses for
the indicated fields.  The soft bilinear coupling $B$ is defined such
that at tree level the mass of the CP-odd MSSM Higgs boson reads
$m_A^2 = 2 B \mu / \sin 2\beta$.  All off-diagonal couplings and
scalar masses are set to zero, as are the first and second generation
Yukawa couplings and soft trilinears.  Although we should stress that
making this choice will lead to a spectrum which is in conflict with
the LHC limits, doing so ensures that fine tuning due to the other
parameters is small, so that we avoid washing out the fine tuning
impact of the stops when the tuning is small\footnote{For models in
  which the spectrum is heavier, when the stop masses are small the
  fine tuning reaches a lower bound imposed by other heavier
  parameters.} as can be the case when the stop masses are less than 1
TeV.  Note that the Higgs mass is also allowed to vary in this scan,
as shown in the right panel of Fig.~\ref{Fig:mssmstoptuning}.  This
illustrates the tuning problem which people have been worrying about
since the discovery of the 125 GeV Higgs boson as we see that raising
the stop masses is also pushing up the Higgs mass, meaning that
heavier Higgs masses require more fine tuning.  However for a low
value of the UV scale this tuning is not so severe unless the stops
are very heavy, and a 125 GeV Higgs can be obtained without much
tuning in this unrealistic case where we have minimized other sources
of tuning.  On the other hand the tuning becomes more severe as we
increase the cutoff such that for $M_X = 10^{16}$ GeV a lightest stop
mass of $1-3$ TeV can result in a fine tuning of $\approx 100 -
1000$ and the minimum tuning we find\footnote{Note that in the
  calculation of the Higgs mass there is a significant theoretical
  error, even with leading two-loop corrections, which should be
  considered when thinking about what the results imply for the
  minimum fine tuning in the model consistent with the recent
  discovery of a 125 GeV Higgs.} for a $125$ GeV Higgs is $\approx
200$, as shown in Fig.~\ref{Fig:mssmstoptuning}.

@@

MSSM benchmark BM2 (dash-dotted in top panel, individual sensitivities in
bottom left panel) is designed to be similar to E$_6$SSM BM2, for a reasonable
comparison.  However from the individual sensitivities one can see that the
behaviour is quite similar to MSSM BM1, though in this case
$\Delta_{m_{H_u}^2}$ becomes the largest tuning at a higher $M_X$ and does not
reach such large values, since more of the tuning is from the mixing in this
case.

@@

Prior to stringent experimental constraints on the mass of the
lightest Higgs boson and squarks in supersymmetric models, a simple
picture of a natural SUSY model emerged from theoretical reasoning,
with soft masses set to similar values at the GUT scale through local
gravity interactions with the hidden sector.  Through the use of
renormalization group running one can then see that at the EW scale
the stops enter the EWSB condition for $M_Z$, therefore it was
expected that these masses should not be much bigger than $100$ GeV.
However to disturb this elegant picture first LEP placed constraints
on the Higgs mass, requiring it to be above 114.4 GeV
\cite{Barate:2003sz, Schael:2006cr}, which already introduced
significant tuning for constrained models, since heavy stops are
required to raise the lightest Higgs mass above it's tree level
bound. Then recently this problem got much worse since the LHC measured the
Higgs mass to be around $125$ GeV.

@@

In contrast we find that in $U(1)$ extensions of the MSSM there is an
additional way to constrain the naturalness of the models, which is
through the $Z^\prime$ mass limit.  We find when we impose a low
cutoff of $20$ TeV for setting the soft masses, the lowest tuning in
the E$_6$SSM compatible with a $Z^\prime$ mass of $2.5$ TeV was
$\Delta \approx 121$, while if the LHC run II can place a limit of
$4.5$ TeV on $M_Z^\prime$ then the tuning would be approximately $394$.
By comparison the current situation in the MSSM only requires a tuning
of around $38$.  This should be interpreted as saying that in
the most conservative limits one can place on naturalness in these
models, the tuning in the E$_6$SSM is worse.  However if there are no
charginos below $700$ GeV then the situation in the two models would
be the same.

@@

When this trace is non-zero it will also induce gauge kinetic mixing
to be generated during RGE evolution and this is the case in the
models we consider here\footnote{In $E_6$ inspired models with only
complete $27$-plet matter multiplets this trace would vanish.
However since we assume some incomplete multiplets so that our
models are consistent with gauge coupling unification this trace
doesn't vanish.}.  However when these models are evolved down from
the GUT scale the radiatively generated gauge kinetic mixing gives an
off-diagonal gauge coupling, $g_{11}$, of just $\approx 0.02$ at the
electroweak scale \cite{King:2005jy} and so it does not play a large
role.  At the same time if gauge kinetic mixing is included the RGE
expressions become very large and unmanageable, so we neglected the
gauge kinetic mixing by setting the SARAH flag \texttt{NoU1Mixing} to true.

@@

& {} - 3 \left ( 3 y_t^4 + 3 y_b^4 + 2 y_t^2 y_b^2 + y_\tau^4 \right ) +

@@

\right ) + 4 y_t^2 y_b^2 \left ( m_{H_u}^2 + m_{H_d}^2 + 2 m_{Q_3}^2 +
